The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-09-18 dit-paper_ieee.tex
Rewritten FPDIT existence condition and construction

		As in the synchronous case, the problem of existence of the FPDIT
		can be written as a system of modular equations:
		\begin{equation}
			\begin{array}{l}
				\exists ? \; t_d \; s.t. \; \forall \tau_i \in \tau, \exists a_i \in [D_i,
				T_i] \; :\\
				\left\{
					\begin{array}{l}
						t_d > O_i \\
						t_d - O_i \equiv a_i \; (mod \;	T_i)
						\\
					\end{array}
				\right.
			\end{array}
		\end{equation}

		This system is guaranteed to have at least one solution modulo $H$ for given
		values of $a_i$ if the following condition is respected:
		%Review 3 : use ref to justify this formula
		\begin{equation}
			a_i + O_i \equiv a_j + O_j \; (\operatorname{mod} \; \operatorname{gcd}(T_i, T_j)) \; \forall i,j
		\end{equation}
		which means that systems where no such values of $a_i$ exists do not have a FPDIT.
		We now present an example of such a system.\\

		\begin{center}
			\begin{tabular}{|r|c|c|c|c|}
				\hline
							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
			\end{tabular}
		\end{center}

		Here, the possible values of $a_1$ are 3 and 4, and the possible values
		of $a_2$ are 7 and 8. For a FPDIT to exist, one of the following
		conditions must be fulfilled:
		\[
			\left[
			\begin{array}{c}
				3 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
				3 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
				4 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
				4 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
			\end{array}
			\right.
		\]
		However neither is, which means the system has
		no FPDIT (and no other DIT than $t=0$).

	\subsection{Solving the modular system}
		The existence conditions of	Section~\ref{sct:FPDITexist} can directly be used to calculate the value of
		the FPDIT (if any), because for a given set of $a_i$ values the problem of
		finding the corresponding DIT is an instance of the Chinese Remainder Problem
		(CRP), which is easily solved using Gauss's CRP solving method for non-coprime
		moduli (Algorithm~\ref{alg:algoCRP}) \cite{gauss1965disquisitiones,knuth1969art}.
		(note that $x^{-1} \; (mod \; y)$ means the modular inverse of $x$ in the naturals modulo $y$)
		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $\forall \tau_i \in \tau : a_i \in [D_i,T_i]$
				\REQUIRE $\forall \tau_i \in \tau : n_i = T_i$
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE Modify the $n_i$ and $a_i$ by prime decomposition so \\ that the
				$n_i$ become pairwise coprime

				\FORALL{$\tau_i \in \tau$}
					\STATE $Mj \leftarrow \frac{H}{n_i}$
					\STATE $invMj \leftarrow Mj^{-1} \; (mod \; n_i)$
					\STATE $e_i \leftarrow Mj \cdot invMj$
				\ENDFOR
				\RETURN $\sum\limits_{\tau_i \in \tau}{a_i \cdot e_i} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}

		%Review 3 : explain Gauss algorithm

		However this is very inefficient as there are exponentially many different
		combinations of $a_i$ values in the number of tasks, and resolving
		exponentially many CRP is prohibitively slow. A faster method for finding
		the FPDIT consists in merging all these CRP instances into one problem
		(finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the
		intermediate results are shared between instances.

@@

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $\forall \tau_i \in \tau: A_i$ is the set of possible $a_i$ values
				for $\tau_i$
				\REQUIRE $\forall \tau_i \in \tau : n_i = T_i$
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE Modify the $n_i$ and $a_i$ by prime decomposition so \\ that the
				$n_i$ become pairwise coprime

				\STATE $sumchunks \leftarrow \text{empty list}$
				\FORALL{$\tau_i \in \tau$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mj \leftarrow \frac{H}{T_i}$
					\STATE $invMj \leftarrow Mj^{-1} (mod \; T_i)$
					\FORALL{$a_i \in A_i$}
						\STATE append $(a_i \cdot Mj \cdot invMj)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[i] \leftarrow taskTerms$
				\ENDFOR

>>>>

		As in the synchronous case, the problem of existence of the FPDIT
		can be written as a system of modular equations:
		\begin{equation}
			\begin{array}{l}
				\exists ? \; t_d \; s.t. \; \forall \tau_i \in \tau, \exists a_i \in [D_i,
				T_i] \; :\\
				\left\{
					\begin{array}{l}
						t_d > O_i \\
						t_d - O_i \equiv a_i \; (mod \;	T_i)
						\\
					\end{array}
				\right.
			\end{array}
		\end{equation}
		This system is equivalent to an instance of the Chinese Remainder Problem
		(CRP) where the goal is to find the value of $t_d$ such that it is congruent
		to $a_i+O_i$ modulo $T_i$ for all tasks $\tau_i$ in $\tau$. We define $x_i =
		a_i+O_i$ and $n_i = T_i$ for clarity in the context of the CRP. This problem
		is guaranteed to have at least one solution modulo $H$ for given values of
		$a_i$ if Condition~\ref{eq:CRPcond} is respected
		\cite{knuth1969art}:
		\begin{equation}
			\label{eq:CRPcond}
			x_i \equiv x_j \; (\operatorname{mod} \; \operatorname{gcd}(n_i, n_j))
			\; \forall \tau_i,\tau_j \in \tau
		\end{equation}
		Condition~\ref{eq:CRPcond} means that DITs might still exist even though in
		the general case task periods will not necessarily be pairwise coprime.
% 		We now present an example of such a system.\\
% 
% 		\begin{center}
% 			\begin{tabular}{|r|c|c|c|c|}
% 				\hline
% 							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
% 				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
% 				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
% 			\end{tabular}
% 		\end{center}
% 
% 		Here, the possible values of $a_1$ are 3 and 4, and the possible values
% 		of $a_2$ are 7 and 8. For a FPDIT to exist, one of the following
% 		conditions must be fulfilled:
% 		\[
% 			\left[
% 			\begin{array}{c}
% 				3 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				3 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 			\end{array}
% 			\right.
% 		\]
% 		However neither is, which means the system has
% 		no FPDIT (and no other DIT than $t=0$).

	\subsection{Solving the modular system}
		A sketch of the construction algorithm for $t_d$ can be found in
		Algorithm~\ref{alg:algoCRP} (note that $x^{-1} \; (mod \; y)$ means the
		modular inverse of $x$ in the naturals modulo $y$). In this paper we
		concentrate on the latter part of the algorithm (after the modifications to
		correct non-coprime periods, readers interested in the details of these
		modifications should refer to \cite{knuth1969art}).		
		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $\forall \tau_i \in \tau : a_i \in [D_i,T_i]$
				\REQUIRE $\forall \tau_i \in \tau : x_i = a_i+O_i$
				\REQUIRE $\forall \tau_i \in \tau : n_i = T_i$
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE Modify the $n_i$ and $x_i$ by prime decomposition so \\ that the
				$n_i$ become pairwise coprime

				\FORALL{$\tau_i \in \tau$}
					\STATE $Mj \leftarrow \frac{H}{n_i}$
					\STATE $invMj \leftarrow Mj^{-1} \; (mod \; n_i)$
					\STATE $e_i \leftarrow Mj \cdot invMj$
				\ENDFOR
				\RETURN $\sum\limits_{\tau_i \in \tau}{x_i \cdot e_i} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}
		Algorithm~\ref{alg:algoCRP} finds one DIT given one value of $a_i$ per task.
		However using it in practice is very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks, and resolving exponentially many CRP instances is prohibitively slow.
		A faster method for finding the FPDIT consists in merging all these CRP
		instances into one problem (finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the intermediate
		results are shared between instances.

@@

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $\forall \tau_i \in \tau: X_i$ is the set of possible $a_i+O_i
				\; \forall a_i \in [D_i,T_i]$ values for $\tau_i$
				\REQUIRE $\forall \tau_i \in \tau : n_i = T_i$
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE Modify the $n_i$ and the content of the $X_i$ by prime
				\\ decomposition so that the $n_i$ become pairwise coprime

				\STATE $sumchunks \leftarrow \text{empty list}$
				\FORALL{$\tau_i \in \tau$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mj \leftarrow \frac{H}{n_i}$
					\STATE $invMj \leftarrow Mj^{-1} (mod \; n_i)$
					\FORALL{$x_i \in X_i$}
						\STATE append $(x_i \cdot Mj \cdot invMj)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[i] \leftarrow taskTerms$
				\ENDFOR
